---
title: "Shark sPAT Tag Maps"
author: "Jennifer Stahl"
date: "August 12, 2019"
output:
  pdf_document: default
  html_document: default
---
##Load packages
Choose the California cran mirror "51" to install packages. 
Install ggmap to download google maps, ggplot2 to plot google maps with points, lines, and polygons, etc., and tidyverse to glimpse data.
```{r message=FALSE, warning=FALSE}
#chooseCRANmirror(ind=51) 
#install.packages("ggmap") 
#install.packages("ggplot2") 
#install.packages("tidyverse") 
#install.packages("devtools")
#devtools::install_github("dkahle/ggmap")
#install.packages ("rgdal") #package to read shapefile.
#install.packages("ggsn")#package for north arrow.
#install.packages("cowplot")#cowplot is an add-in to ggplot
#install.packages("ggspatial")

library(ggmap)
library(ggplot2)
library(tidyverse)
#library(devtools)
#library(rgdal) 
#library(ggsn)
library(ggspatial)
library(cowplot) #for insert map
```

## Get base maps
###Register google key 
```{r include=FALSE, warning=FALSE}
map_key <-read.csv(file="C:/R_git_stuff/Mobula-tag-maps/Map_key.csv", header=TRUE, sep=",")
register_google(key = map_key$Key) 
```

### Import map of Pacific centered on Hawaii.
```{r import Pacific map}
Pacgooglemap <- get_map(location = c(lon = -155, lat = 20),
                       source   = "google",
                       maptype  = "satellite", #can also do "terrain" for a simplier map.
                       zoom     = 2)# This is the zoom parameter, larger number zooms in more.

#zoomed in more
Pacgooglemapzoom <- get_map(location = c(lon = -155, lat = 20),
                       source   = "google",
                       maptype  = "satellite", 
                       zoom     = 3)

Pacgooglemapzoom_4 <- get_map(location = c(lon = -155, lat = 20),
                       source   = "google",
                       maptype  = "satellite", 
                       zoom     = 4)

```
###Display google map using ggmap, which uses grid graphics.
```{r display Pacific map}
Pacbasemap<- ggmap(Pacgooglemap)
Pacbasemap

Pacbasemapzoom<- ggmap(Pacgooglemapzoom)
Pacbasemapzoom

Pacbasemapzoom_4<- ggmap(Pacgooglemapzoom_4)
Pacbasemapzoom_4

```

##Import shark data
```{r import data}
sharks <-read.csv(file="C:/R_git_stuff/Shark-tag-maps/LL_TAG_DATA_GOOD_Nov14_2022.csv", header=TRUE, sep=",")
#head(sharks)
#glimpse(sharks)
#dim(sharks)
```

##Map shark tag data.
###First map tag releases and recoveries as point data and look for outliers.
```{r map tag points}
tag_rel <- Pacbasemapzoom + geom_point(data = sharks, aes(DEPLOY_LON_DD, DEPLOY_LAT_DD), colour = "red")
tag_rel 
tag_rec <- Pacbasemapzoom + geom_point(data = sharks, aes(POPOFF_LON_DD, POPOFF_LAT_DD), colour = "yellow")
tag_rec 
#Tags in eastern latitude will not map, so need to run below code and rerun.
#Also anything with no popoff location will not map.
```
Ggplot maps on a 0-360 scale so does not change to positive 180 when crossing the antimeridian. Couldn't figure out how to fix this.
Instead converted tags across the antimeridian to a 0-360 longitude, so they will display on the map. 
```{r correct longitudes}
sharks$POPOFF_LON_DD[which(sharks$POPOFF_LON_DD>0)] <- -360 + (sharks$POPOFF_LON_DD[which(sharks$POPOFF_LON_DD>0)])
```

###Map tag paths from release to recovery.
```{r map tagpaths}
tagpaths <- Pacbasemapzoom + geom_segment(aes(x = DEPLOY_LON_DD, y = DEPLOY_LAT_DD, xend = POPOFF_LON_DD, yend = POPOFF_LAT_DD, colour = SPECIES), arrow = arrow(angle = 30, length = unit(0.1, "cm"), ends = "last", type = "open"), data = sharks)
tagpaths
ggsave("Shark tag paths.jpg") 
```

******************************************************************
##Reference map 
### Purpose to create a reference map if zoomed in on main map.

###Get bounding box values for reference map.
####Note - Lat is y and Lon is x.
```{r bounding box}
Paczoombb <-attr(Pacgooglemapzoom, "bb") #this only extracts the bounding box for the google map not the map transformed with ggmap.
Paczoombb 
```

###Create a data frame of the 4 polylines that define bounding box of reference map.
```{r bb df}
Pacmapzoom_bblines    <- data.frame(x1 = c(-211, -98.7, -98.7, -211),
                                y1 = c(-33.8, -33.8, 60.6, 60.6),
                                x2 = c(-98.7, -98.7, -211, -211),
                                y2 = c(-33.8, 60.6, 60.6, -33.8))
```

###Add the bounding box layer to the map of tagpaths
```{r map bb}
Paczoom_insert_map <-tagpaths + geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2, colour = "red"), 
                               show.legend = FALSE, data = Pacmapzoom_bblines)
Paczoom_insert_map
```

###Remove the x and y axes from the Pacific reference map.
```{r remove axes}
Paczoom_insert_map_nobackground <- Paczoom_insert_map +
      theme(plot.background =
      element_rect(fill = NA, linetype = 1,
      size = 0.01, colour = "black"),
      axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.position="none")
```
### Display the Pacific reference map with the final Pac map map.
```{r reference map}
Pac_tag_map_final <- ggdraw() +
  draw_plot(Pac_tag_map_arrow_and_bar, 0, 0, 1, 1) +
  draw_plot(Paczoom_insert_map_nobackground, 0.2,0.7,0.25,0.25)
Pac_tag_map_final
ggsave("Pac_tag_map_final.jpg") 
```

******************************************************************
#Map elements

##Scalebar
###Issues with below map elements which is from old code and some of the old packages are no longer installing.
```{r message=FALSE, warning=FALSE}
install.packages("grid") #Need to install grid package before ggsn.
library(grid)
#install.packages("ggsn") #Use the development version of ggsn intead in the next line.
devtools::install_github('oswaldosantos/ggsn')#Development version of ggsn.If doesn't update correctly, then can add force = TRUE from code in next line.
#devtools::install_github('oswaldosantos/ggsn', force = TRUE)
library(ggsn)
```

###Will need bounding box for each google base map for scalebar function.
Note that Lat is y and Lon is x.
```{r bounding box}
Paczoombb <-attr(Pacgooglemapzoom, "bb") #this only extracts the bounding box for the google map not the map transformed with ggmap.
Paczoombb 
```

### Add scalebar layer to map.
Need to indicate you are using the scalebar from the ggsn package (ggsn::scalebar) because there is also a scalebar function in the raster package.The first package you install can mask the function from the other package if both packages are installed.

```{r add scalebar to map}
Pac_tag_map_with_scalebar <- tagpaths + ggsn::scalebar(dist = 2000, dist_unit = "km", transform = TRUE, model = "WGS84",
                           height = 0.02, st.dist = 0.05, st.bottom = TRUE, st.size = 4, st.color = "white", 
                           border.size = 0.05, x.min = -211, y.min = -33.8, x.max = -98.7, y.max = 60.6, 
                           anchor = c(x=-110, y=-22))
Pac_tag_map

```

##North arrow
### the north arrow is displaying but error that it isn't meaningful...
### info on north arrow in ggspatial package: https://rdrr.io/cran/ggspatial/man/annotation_north_arrow.html

```{r}
Pac_tag_map_arrow_and_bar <- tagpaths + geom_spatial_point(aes()) +
  annotation_north_arrow(which_north = "true", location="tr", height = unit(0.75, "cm"),
  width = unit(0.75, "cm"),)


Pac_tag_map_arrow_and_bar
ggsave("Pac_tag_map_arrow_and_bar.jpg") #
```

*******************************************************************
#EEZ 
##The below code is old code to get EEZ layer that no longer works. I think the packages are out of date.

##Add US EEZ layer to base maps.
###Import shapefile for EEZs for entire world.
```{r message=FALSE, warning=FALSE}
#use rgdal package to read shapefile from location on your computer.
eez.world <- readOGR(dsn = "C:/R stuff/Maps_longline", layer = "eez_v11") #Each file that makes up the shapefile layer needs to be in the working directory folder. Will get an error if have within another folder that isn't the working directory. 
```

###Extract specific EEZs from world EEZ shapefile.
Creates a large SpatialPolygonsDataFrame.
Can first determine names of US EEZs in Pacific by filtering .dbf of shapefile in excel. Then can subset for EEZs in R.
If want to just extract HI EEZ can use this code in the future: dat.eez.hi <-eez.world[eez.world@data$Territory1 == "Hawaii",] 
```{r subset US EEZ}
dat.eez.pac <-eez.world[eez.world@data$TERRITORY1 %in% c("Hawaii","American Samoa","Northern Mariana Islands", "Jarvis Island", "Howland and Baker Islands", "Johnston Atoll", "Wake Island", "Palymyra Atoll"),]
```

###Put the EEZ data in a data frame that can be used by ggplot.
```{r create EEZ data frame with subset}
dat.eez.pac_df <-fortify(dat.eez.pac)
dim(dat.eez.pac_df)
```

###Add EEZ layer to the base maps.
```{r map EEZ}
Pacbase_eez <- Pacbasemapzoom +  geom_path(data = dat.eez.pac_df,
                                     aes(x = long, y = lat, group = group),
                                     colour = "white", size = 0.5)
Pacbase_eez 
```